---
title: "08-10-2020-TT-Avatar"
author: "Brett Moxham"
date: "10/08/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Let's load the data, and our libraries. Our data comes from the [Tidy Tuesday weekly data challenge.](https://github.com/rfordatascience/tidytuesday/blob/master/data/2020/2020-08-11/readme.md) 


```{r}
library(tidyverse)


avatar <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-08-11/avatar.csv')
scene_description <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-08-11/scene_description.csv')

```


## EDA

```{r}
skimr::skim(avatar)

skimr::skim(scene_description)


avatar %>% 
  filter(is.na(imdb_rating)) %>% 
  view()
```


Avatar is missing 1 `imdb_rating` for "The Seige of North, Part 2". It probably didn't parse correctly. We can find that rating on the [IMDB Website](https://www.imdb.com/title/tt10512638/). Let's add in the missing rating to complete our dataset. The missing rating is *9.7*. We can use the `chapter_num` and `book_num` to make sure we add our data correctly.

```{r}
avatar %>% 
      mutate(
        imdb_rating = case_when( book_num == 1 & chapter_num == 20 ~ 9.7,
                                 TRUE ~ imdb_rating
          
        )
      ) -> avatar

skimr::skim(avatar)
```

We've now have a complete set of `imdb_ratings`.

## Visualization


```{r}
distinct_avatar <- avatar %>% 
  select(book, book_num, chapter, chapter_num, imdb_rating) %>% 
  distinct(book, book_num, chapter, chapter_num, imdb_rating)
```


```{r}
plot_tile <- distinct_avatar %>% 
  ggplot(aes(
    x =fct_reorder(book, book_num),
    y = chapter_num,
    label = imdb_rating
  )) +
  geom_tile(aes(fill = imdb_rating))+
  geom_text(aes(x = book, y = chapter_num, label = imdb_rating ), alpha = 0.2, size = 4, fontface = "bold", color = "black")+
  scale_y_reverse(breaks = round(seq(min(distinct_avatar$chapter_num), max(distinct_avatar$chapter_num), by = 1),1))+
  scale_fill_distiller(palette ="Reds")+
  bmox::theme_bmox()+
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    plot.caption = element_text(size = 7)
  ) +
  labs(
    title = "Avatar: The Last Airbender",
    subtitle = "Episode Ratings by Book (Season)",
    x = NULL,
    y = "Episode Number",
    fill = "IMDB Rating",
    caption = "@bmoxO9 | #TidyTuesday 
Source: https://tinyurl.com/y65yx8l7"
  )
```

```{r}
ggsave("Images/Avatar/avatar_heatmap.png", plot_tile, height = 6, width = 6)
```

